---
title: "Group 3 Data Cleaning and Analysis"
author: "Anum, Blagoja, Hannah, Nolan" 
date: "Final Report Due Monday, March 13, 2023"
output: 
  github_document: 
    toc: true
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
# This setup chunk sets global defaults and includes the tidyverse packages
# The option include = FALSE prevents warnings and messages from printing to your report.
# Change the next option to echo = TRUE to hide code chunks by default 
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

# Introduction

Project Proposal:
<https://docs.google.com/document/d/12Ly0ewjRK-uhTVQvT38E0vjMvc8uvwr67GpmBKwA2ak/edit?usp=sharing>

In this project, we are going to be using the two datasets that we found
on Kaggle: "Highest Holywood Grossing Movies.csv" and
"imdb_top_1000.csv". We would be renaming them as
'Highest_Holywood_Grossing_Movies' and 'imdb_top_1000' dataframes to
work with in our project.

## Information about each datasets from the project

### Highest Holywood Grossing Movies.csv dataset

Data frame Highest_Holywood_Grossing_Movies:  
Rows: 918.  
Columns: 11.  
Variables (Column Names): Title, Movie Info, Distributor, Release Date,
Domestic Sales, International Sales, World Sales, Genre, Movie Runtime,
and License.

### imdb_top_1000.csv dataset

Data frame imdb_top_1000:  
Columns: 16  
Rows: 1,000  
Variables (Column Names): Poster_Link, Series_Title, Released_Year, Certificate, Runtime,
Genre, IMDB_Rating, Overview, Meta_score, Director, Star1, Star2, Star3,
Star4, No_of_votes, and Gross.

## Target variables of Data Science Project

The variables that we are going to work with from the two datasets are
Title, Release Year, Runtime, Genre, Domestic Gross, World Gross,
Distributor, and perhaps others.  

In terms of variables we will create, we are interested in perhaps
trying to create a new categorical variable of the gender of different
lead actors, separating the different genres movies are classified in,
and other depending on which questions we choose to focus on.  

### Call to the original two datasets:

```{r data, echo = TRUE}
gross <- read_csv("data/Highest Holywood Grossing Movies.csv")
imdb <- read_csv("data/imdb_top_1000.csv")

glimpse(gross)
glimpse(imdb)


```

### Tidying of the IMDB dataset:

```{r imdb_tidy, echo = TRUE}
#imdb2: clean names, rename some variables, select vars, add wrong data point.
imdb2 <- imdb %>%
clean_names() %>%
rename(title = series_title) %>%
select(-poster_link)
# there's no reason to use Base R for this
imdb2$released_year <- replace(imdb2$released_year, 967, "1995")

glimpse(imdb2)

```

### Tidying of the Top Grossing Movies dataset:

```{r gross_tidy, echo = TRUE}
#gross2: clean names, rename vars, fix runtime variable
gross2 <- gross %>%
  clean_names() %>%
  rename(runtime = movie_runtime) %>%
  mutate(hours = word(runtime, 1), minutes = word(runtime, 3)) %>%
  select(-runtime) %>%
  mutate(hours = parse_number(hours)) %>%
  mutate(minutes = parse_number(minutes)) %>%
  mutate(hours = hours * 60) %>%
  mutate(runtime = hours + minutes) %>%
  select(-hours, -minutes)
  
#gross3: tidy title variable.
gross3 <- gross2 %>%
  separate(title, into= c("movie_title", "extra_year"), sep="\\(") %>%
  select(-extra_year)

#gross4: create action/comedy/drama var.
gross4 <- gross3 %>%
  mutate(action = if_else(str_detect(genre, "Action")==TRUE, "Y" , "N"),
         comedy = if_else(str_detect(genre, "Comedy")==TRUE, "Y" , "N"),
         drama = if_else(str_detect(genre, "Drama")==TRUE, "Y" , "N"))


#gross5: create combo var from action/comedy/drama vars.
gross5 <- gross4 %>%
  mutate(combo = if_else(str_detect(action, "Y") & str_detect(comedy, "Y")& str_detect(drama, "Y"), "Action, Comedy, and Drama" ,
                 if_else(str_detect(action, "Y") & str_detect(comedy,"N") & str_detect(drama,"N") , "Action only" ,
                 if_else(str_detect(action, "N") & str_detect(comedy,"Y") & str_detect(drama,"N") , "Comedy only" ,
                 if_else(str_detect(action, "N") & str_detect(comedy,"N") & str_detect(drama,"Y") , "Drama only"  ,
                 if_else(str_detect(action, "N") & str_detect(comedy,"Y") & str_detect(drama,"Y") , "Drama and Comedy" ,
                 if_else(str_detect(action, "Y") & str_detect(comedy,"N") & str_detect(drama,"Y") , "Drama and Action" ,
                 if_else(str_detect(action, "Y") & str_detect(comedy,"N") & str_detect(drama,"Y") , "Drama and Action" ,
                 "Neither"))))))))

#gross6: rename/edit variables, trim title var.
gross6 <- gross5 %>%
  select(-x1) %>%
  rename(title = movie_title)

gross6$title <- str_trim(gross6$title)

glimpse(gross6)

```

### Creation of merge dataset:

```{r merge_tidy, echo = TRUE}
#Combine gross6 and imdb2 into the merge dataset, and rename some vars, tidy dates, and create month variable.
merge <- inner_join(gross6, imdb2, by="title") %>%
  select(title, movie_info, distributor, release_date, released_year, domestic_sales_in, world_sales_in, license, runtime.x, combo, imdb_rating,
         meta_score, director, star1, no_of_votes) %>%
  # Can use 1 rename command to handle all of these.
  rename(dom_sales = domestic_sales_in) %>%
  rename(world_sales = world_sales_in) %>%
  rename(runtime = runtime.x) %>%
  rename(imdb_no_votes = no_of_votes) %>%
  rename(release_year = released_year) %>%
  mutate(release_date = as.Date(release_date, "%B %d, %Y"))

merge$month <- format(as.Date(merge$release_date, format="%d/%m/%Y"),"%m")
merge$month <- as.factor(merge$month)

glimpse(merge)

```

## Data Visualization

### Rating:

```{r rating_an, echo = TRUE}

#IMDB rating by meta-score
merge %>%
 ggplot(aes(x = meta_score , y = imdb_rating, color = combo)) +
 geom_jitter(alpha = .5) +
 geom_smooth( se = FALSE) +
 labs(title="Relationship between Meta Score and IMDB ratings", x="Metascore", y="IMDB ratings" , fill = "Genres")


merge %>%
  filter(meta_score>95) %>%
  arrange(desc(meta_score)) %>%
  select(title, meta_score, imdb_rating, director)

#Correlation between imdb rating and runtime?
merge %>%
  ggplot(aes(x=runtime, y=imdb_rating)) +
  geom_jitter() +
  geom_smooth() + 
  labs(title="Runtime by IMDB rating", x="Movie runtime", y="IMDB rating") + 
  theme(plot.title = element_text(hjust = 0.5))

#What is the most common imdb rating?
merge %>%
  ggplot(aes(x=factor(imdb_rating))) +
  geom_bar() + 
  labs(title="Distribution of IMDB ratings", x="IMDB rating", y="Count") + 
  theme(plot.title = element_text(hjust = 0.5))

#What are the films which got the highest IMDB rating?
merge %>%
  filter(imdb_rating >= 9) %>%
  select(title, imdb_rating)


```

### Sales:

```{r sales_an, echo = TRUE}

#compare domestic and world sales
merge %>%
  ggplot(aes(x=dom_sales, y=world_sales)) +
  geom_point() +
  geom_smooth(se=FALSE)+ 
  labs(title="Domestic by World Sales", x="Domestic Sales", y="World Sales") + 
  theme(plot.title = element_text(hjust = 0.5))

#Far right outliers
merge %>%
  select(title, dom_sales, world_sales, release_year) %>%
  arrange(desc(world_sales))%>%
  filter(dom_sales > 6.3e+08) 

#IMDB score and sales
merge %>%
  ggplot(aes(x=dom_sales, y=imdb_rating)) +
  geom_jitter() +
  geom_smooth() + 
  labs(title="Domestic Sales by IMDB rating", y="IMDB rating", x="Domestic sales") + 
  theme(plot.title = element_text(hjust = 0.5))

#metascore and sales
merge %>%
  ggplot(aes(x=dom_sales, y=meta_score)) +
  geom_jitter() +
  geom_smooth() + 
  labs(title="Domestic Sales by Meta score", x="Domestic Sales", y="IMDB rating") + 
  theme(plot.title = element_text(hjust = 0.5))



```

### Distributor:

```{r distributor_an, echo = TRUE}
#List of all different distributors
merge %>%
  select(distributor) %>%
  distinct() %>%
  arrange(distributor)
  

#Sales by Dist
merge %>%
  group_by(distributor) %>%
  select(distributor, dom_sales)%>%
  ggplot() +
  geom_col(aes(x=fct_infreq(distributor), y=dom_sales, fill=distributor), show.legend = FALSE) +
  coord_flip() +
  labs(title="Sales by Distributor", y="Total Domestic Sales", x="Distributor") + 
  theme(plot.title = element_text(hjust = 0.5))

#Why does Disney have so much more? (ans: has a lot of the top earning movies/has highest quantity in list)
merge %>%
  group_by(distributor) %>%
  filter(dom_sales >= 504481165) %>%
  select(title, distributor, dom_sales) %>%
  arrange(desc(dom_sales))

#What is the count of each distributor in the combined dataset?
merge %>%
  group_by(distributor) %>%
  summarise(n=n()) %>%
  arrange(desc(n)) %>%
  slice_head(n=10)

#Sale by distributor
merge %>%
  ggplot(aes(x=distributor, y= log(dom_sales)))+
  geom_boxplot()+
  coord_flip() + 
  labs(title="Distributor by Domestic Sales", x="Distributor", y="Domestic Sales (log)") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### Starring actor and director:

```{r director_and_star_an, echo = TRUE}
#total sales by star1
merge %>%
  group_by(star1) %>%
  summarise(total_sales_star = sum(dom_sales)) %>%
  select(star1, total_sales_star) %>%
  slice_head(n=20) %>% 
  ggplot() +
  geom_col(aes(y=reorder(star1, total_sales_star), x=total_sales_star, 
           fill=reorder(star1, total_sales_star)),
           show.legend = FALSE) + 
  labs(title="Total Domestic Sales of Movie by Star1", x="Total Sales of Movies with Star1", y="Star1") + 
  theme(plot.title = element_text(hjust = 0.5))

#Count of most common star
merge %>%
  group_by(star1) %>%
  summarise(n=n()) %>%
  arrange(desc(n)) %>%
  slice_head(n=10)

#Most common star1, domestic sales, and IMDB rating
merge %>%
 filter( star1 %in% c("Tom Hanks", "Leonardo DiCaprio", "Daniel Radcliffe", "Tom Cruise", "Daniel Craig", "Joe Russo", "Matt Damon", "Christian Bale", "Denzel Washington", "Elijah Wood")) %>%
 ggplot(aes(x = factor(imdb_rating) , y = dom_sales)) +
 geom_col()+
 facet_wrap(~star1) +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
 labs(title="Relationship between IMDB rating and sales base on star1", x="IMDB rating" , y = "Sales")


#Total sales by director
merge %>%
  group_by(director) %>%
  summarise(total_sales_dir = sum(dom_sales)) %>%
  select(director, total_sales_dir) %>%
  slice_head(n=20) %>% 
  ggplot() +
  geom_col(aes(y=reorder(director, total_sales_dir), x=total_sales_dir, 
           fill=reorder(director, total_sales_dir)),
           show.legend = FALSE) + 
  labs(title="Total Domestic Sales of Movie by Director", x="Total Sales of Movies with Director", y="Director") + 
  theme(plot.title = element_text(hjust = 0.5))


#Count of most common director
merge %>%
  group_by(director) %>%
  summarise(n=n()) %>%
  arrange(desc(n)) %>%
  slice_head(n=10) # Use pull() here to get the list of names to save for use below.
# Then you could pipe it in.


#Most common director, domestic sales, and IMDB rating
merge %>%
 filter( director %in% c("Anthony Russo" ,"Christopher Nolan","Clint Eastwood", "David Fincher","James Cameron","Peter Jackson","Ridley Scott" , "Robert Zemeckis", "Sam Mendes","Steven Spielberg")) %>%
 ggplot(aes(x = factor(imdb_rating) , y = dom_sales)) +
 geom_col()+
 facet_wrap(~director) +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
 labs(title="Relationship between IMDB rating and sales base on director", x="IMDB rating" , y = "Sales")

merge %>%
  filter( director %in% c("Anthony Russo" ,"Christopher Nolan","Clint Eastwood", "David Fincher","James Cameron","Peter   Jackson","Ridley Scott" , "Robert Zemeckis", "Sam Mendes","Steven Spielberg")) %>%
  ggplot(aes(x=runtime, y=director)) +
  geom_boxplot() + 
  labs(title="Runtime by Director", x="Runtime", y="Director") + 
  theme(plot.title = element_text(hjust = 0.5))


```

### Release month:

```{r month_an, echo = TRUE}

#Graph dom profit by month variable
merge %>%
  filter(!is.na(month)) %>%
  ggplot(aes(x=month, y=dom_sales)) +
  geom_col(aes(fill=month), show.legend = FALSE) + 
  labs(title="Domestic Profit by Month", x="Month", y="Domestic Sales") + 
  theme(plot.title = element_text(hjust = 0.5))

  
#graph world profit by season variable (not very different!)
merge %>%
  filter(!is.na(month)) %>%
  ggplot(aes(x=month, y=world_sales)) +
  geom_col(aes(fill=month), show.legend = FALSE) + 
  labs(title="World Profit by Month", x="Month", y="World Sales") + 
  theme(plot.title = element_text(hjust = 0.5))

#Number of films by month
merge %>%
  filter(!is.na(month)) %>%
  ggplot() +
  geom_bar(aes(x=month, fill=month), show.legend = FALSE) + 
  labs(title="Number of Films per Month", x="Month", y="Count") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### Runtime:

```{r runtime_an, echo = TRUE}

#Runtime and domestic sales
merge %>%
  ggplot(aes(x=runtime, y=dom_sales)) +
  geom_point() +
  geom_smooth() + 
  labs(title="Runtime by Domestic Sales", x="Runtime", y="Domestic Sales") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### Release year:

```{r year_an, echo = TRUE}

#Sales by Year (inflation? popularity of seeing movies?)
merge %>%
  ggplot(aes(x=factor(release_year), y=dom_sales)) +
  geom_jitter() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title="Sales by Release Year", x="Release Year", y="Domestic Sales") + 
  theme(plot.title = element_text(hjust = 0.5))

#Release year by combo / popular genres
merge %>%
 ggplot(aes(x = release_year, fill = combo)) +
 geom_bar() +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
 labs(title="Popular Genres among different years", x="Year", fill = "Genres") +
 geom_bar(width = 0.5) + 
  labs(title="Popular Genres by Year", x="Release Year", y="Type(s) of Genre") + 
  theme(plot.title = element_text(hjust = 0.5))


```

### Genre:

```{r genre_an, echo = TRUE}

#Combo and IMDB rating
ggplot(merge, aes(x=imdb_rating, y=combo)) + 
  geom_boxplot()+
  labs(title= "Genre and IMDB rating", x= "imdb rating", y= "combo") + 
  theme(plot.title = element_text(hjust = 0.5))

#IMDB rating with domestic sales and genre
# Nice graphic, just need to fill in the empty spot at 9.1 so that the scale is numeric

merge %>%
ggplot(aes(x = factor(imdb_rating) , y = dom_sales , fill = combo)) +
geom_col() +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
labs(title="IMDB Ratings and Domestic Sales in regards to different genres", x="IMDB Rating", y="Domestic Sales" , fill = "Genres") +
 theme(plot.title = element_text(hjust = 0.5))

merge %>%
ggplot(aes(x = runtime , y = dom_sales))+
 geom_point()+
 geom_smooth(method = "lm" , se = FALSE)+
 facet_wrap(~combo) +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
 labs(title="Relationships between length of the movies and sales based on their genre", x=" Movie Length", y="Sales") +
 theme(plot.title = element_text(hjust = 0.025))
```
